---
title: "Project 1"
author: "Ella Acosta"
date: "9/30/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
setwd("~/Desktop/DATS 6101/Project 1")

df <- read.csv("ShanghaiPM.csv")
```

```{r, echo=FALSE}
str(df)
names(df)
# remove index column
df <- df[,-1]
# omit rows with NAs
df <- na.omit(df)

# check factor levels and ranges for numeric and interger variables
range(df$year) # data collected from 2013-2015
df$year <- as.integer(df$year)
range(df$hour) # hour is military time
range(df$season) # 1-4, change to factor spring, summer, fall, winter
```



```{r, echo=FALSE}
#rename month variable
colnames(df)[2] <- "rawMonth"
#create factor variable
df$month[df$rawMonth == 1] <- "Jan"
df$month[df$rawMonth == 2] <- "Feb"
df$month[df$rawMonth == 3] <- "Mar"
df$month[df$rawMonth == 4] <- "Apr"
df$month[df$rawMonth == 5] <- "May"
df$month[df$rawMonth == 6] <- "Jun"
df$month[df$rawMonth == 7] <- "Jul"
df$month[df$rawMonth == 8] <- "Aug"
df$month[df$rawMonth == 9] <- "Sep"
df$month[df$rawMonth == 10] <- "Oct"
df$month[df$rawMonth == 11] <- "Nov"
df$month[df$rawMonth == 12] <- "Dec"

df$month <- as.factor(df$month)
```

Variables
PM: pm2.5 concentration (ug/m^3)
DEWP: dew point (celsius)
TEMP: temperature (celsius)
HUMI: humidity (%)
PRES: pressure (hPa)
cbwd: combined wind direction 
lws: cumulated wind speed (m/s)
precipitation: hourly precipitation (mm)
Iprec: cumulated precipitation (mm)
```{r, echo=FALSE}
#create new season variable
#season 1 is spring
#season 2 is summer
#season 3 is fall
#season 4 is winter

#rename season variable
colnames(df)[5] <- "rawSeason"
df$season[df$rawSeason == 1] <- "Spring"
df$season[df$rawSeason == 2] <- "Summer"
df$season[df$rawSeason == 3] <- "Fall"
df$season[df$rawSeason == 4] <- "Winter"

df$season <- as.factor(df$season)
```

```{r, echo=FALSE}
#create averaged pm2 column
library(dplyr)
df <- mutate(df, avg_pm2 = (PM_Jingan + PM_US.Post + PM_Xuhui)/3)
```

```{r, echo=FALSE}
# drop raw data variables
df <- df[, -c(2, 5:8)]
names(df)

# reorganize variables
df <- df[,c(1, 13, 12, 2, 3, 14, 4:11)]
```

Exlporatory Data Analysis
```{r, echo=FALSE}
dfSum <- summary(df)

# run boxplots with outliers
```

```{r, echo=FALSE}
# remove outliers

outlierKD <- function(dt, var) { 
     var_name <- eval(substitute(var),eval(dt))
     na1 <- sum(is.na(var_name))
     m1 <- mean(var_name, na.rm = T)
     par(mfrow=c(2, 2), oma=c(0,0,3,0))
     boxplot(var_name, main="With outliers")
     hist(var_name, main="With outliers", xlab=NA, ylab=NA)
     outlier <- boxplot.stats(var_name)$out
     mo <- mean(outlier)
     var_name <- ifelse(var_name %in% outlier, NA, var_name)
     boxplot(var_name, main="Without outliers")
     hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
     title("Outlier Check", outer=TRUE)
     na2 <- sum(is.na(var_name))
     cat("Outliers identified:", na2 - na1, "n")
     cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "n")
     cat("Mean of the outliers:", round(mo, 2), "n")
     m2 <- mean(var_name, na.rm = T)
     cat("Mean without removing outliers:", round(m1, 2), "n")
     cat("Mean if we remove outliers:", round(m2, 2), "n")
     response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
     if(response == "y" | response == "yes"){
          dt[as.character(substitute(var))] <- invisible(var_name)
          assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
          cat("Outliers successfully removed", "n")
          return(invisible(dt))
     } else{
          cat("Nothing changed", "n")
          return(invisible(var_name))
     }
}

# running the outlierKD code below triggers a prompt on the console window (yes/no). Answer there to proceed.
outlierKD(df, df$avg_pm2)
#remove NAs
df <- na.omit(df)
```

Boxplots for PM2 by season for each year
```{r, echo=FALSE}
library(ggplot2)
library(magrittr)
library(grid)
library(gridExtra)

fill_2013 <- "green3"
line_2013 <- "#1F3552"
boxplot2013 <- df %>%
  select(avg_pm2, season, year) %>%
  filter(year == 2013) %>%
  ggplot(., aes(season, avg_pm2)) + 
  geom_boxplot(fill = fill_2013, colour = line_2013) + scale_x_discrete(name = "Season") + scale_y_continuous(name = "Cumulative Particulate\nMatter") + ggtitle("Boxplot of Cumulative Particulate\nMatter by Season in 2013") + theme(plot.title = element_text(hjust = 0.5))

fill_2014 <- "deeppink3"
line_2014 <- "#1F3552"
boxplot2014 <- df %>%
  select(avg_pm2, season, year) %>%
  filter(year == 2014) %>%
  ggplot(., aes(season, avg_pm2)) + 
  geom_boxplot(fill = fill_2014, colour = line_2014) + scale_x_discrete(name = "Season") + scale_y_continuous(name = "Cumulative Particulate\nMatter") + ggtitle("Boxplot of Cumulative Particulate\nMatter by Season in 2014") + theme(plot.title = element_text(hjust = 0.5))

fill_2015 <- "blue3"
line_2015 <- "#1F3552"
boxplot2015 <- df %>%
  select(avg_pm2, season, year) %>%
  filter(year == 2015) %>%
  ggplot(., aes(season, avg_pm2)) + 
  geom_boxplot(fill = fill_2015, colour = line_2015) + scale_x_discrete(name = "Season") + scale_y_continuous(name = "Cumulative Particulate\nMatter") + ggtitle("Boxplot of Cumulative Particulate\nMatter by Season in 2015") + theme(plot.title = element_text(hjust = 0.5))

grid.arrange(boxplot2013, boxplot2014,
             boxplot2015, ncol = 2)
```

Particulate Matter Scatterplots by Continous Explanatory Variables
```{r, echo=FALSE}
by_dew2013 <- df %>%
  select(avg_pm2, DEWP, year, month) %>%
  filter(year == 2013) %>%
  ggplot(., aes(DEWP, avg_pm2)) + geom_point(aes(colour = month)) 

by_humi2013 <- df %>%
  select(avg_pm2, HUMI, year, month) %>%
  filter(year == 2013) %>%
  ggplot(., aes(HUMI, avg_pm2)) + geom_point(aes(colour = month))

by_pres2013 <- df %>%
  select(avg_pm2, PRES, year) %>%
  filter(year == 2013) %>%
  ggplot(., aes(PRES, avg_pm2)) + geom_point(aes(colour = month))

by_temp2013 <- df %>%
  select(avg_pm2, TEMP, year) %>%
  filter(year == 2013) %>%
  ggplot(., aes(TEMP, avg_pm2)) + geom_point(aes(colour = month))

by_precip2013 <- df %>%
  select(avg_pm2, year, precipitation) %>%
  filter(year == 2013) %>%
  ggplot(., aes(precipitation, avg_pm2)) + geom_point(aes(colour = month))

arrange(by_dew2013, by_humi2013, by_pres2013, by_temp2013, by_precip2013, ncol = 2)
```

Aggregate Time Series PM2 Levels by Month
```{r, echo=FALSE}
#aggregate hourly pm2 levels by month
df2013 <- df %>%
  select(avg_pm2, year, month, DEWP, HUMI, PRES, TEMP, precipitation) %>%
  filter(year == 2013) %>%
  group_by(month) %>%
  summarise(., mean_pm2 = mean(avg_pm2), mean_DEWP  = mean(DEWP),
            mean_HUMI = mean(HUMI), mean_PRES = mean(PRES),
            mean_TEMP = mean(TEMP), mean_PREC = mean(precipitation)) 

df2013plot <- df2013 %>%
  ggplot(., aes(month, mean_pm2)) + geom_bar(stat = "identity")

df2014 <- df %>%
  select(avg_pm2, year, month, DEWP, HUMI, PRES, TEMP, precipitation) %>%
  filter(year == 2014) %>%
  group_by(month) %>%
  summarise(., mean_pm2 = mean(avg_pm2), mean_DEWP  = mean(DEWP),
            mean_HUMI = mean(HUMI), mean_PRES = mean(PRES),
            mean_TEMP = mean(TEMP), mean_PREC = mean(precipitation))

df2013plot <- df2013 %>%
  ggplot(., aes(month, mean_pm2)) + geom_bar(stat = "identity")

df2015 <- df %>%
  select(avg_pm2, year, month, DEWP, HUMI, PRES, TEMP, precipitation) %>%
  filter(year == 2015) %>%
  group_by(month) %>%
  summarise(., mean_pm2 = mean(avg_pm2), mean_DEWP  = mean(DEWP),
            mean_HUMI = mean(HUMI), mean_PRES = mean(PRES),
            mean_TEMP = mean(TEMP), mean_PREC = mean(precipitation))
```

```{r, echo=FALSE}
# define levels of granularity 
# stratified graphs


```